---
title: "Model"
author: "Anna Flieder"
date: "10/26/2021"
output: html_document
---

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
library(reshape2)
library(car)
library(deSolve)
library(readxl)
```

```{r}
#probability of someone seeking promotion to next level
#u= fraction of like-gendered individuals in level above
#v= fraction of like-gendered individuals in current level
#lambda= strength of homophilic tendency
e <- exp(1)

P_func <- function(u, v, lambda){
  e <- exp(1)
  denom <- 1 + e^(-lambda*(u-v))
  return(1/denom)
}

#fraction of women in applicant pool
F0_func <- function(u, v, lambda){
  num <- v*P_func(u, v, lambda)
  denom <- num + ((1-v)*(P_func(1-u, 1-v, lambda)))
  return(num/denom)
}

#fraction of women promoted to next level
#b= bias (b>0.5 means women are favored)
F_func <- function(u, v, lambda, b){
  num <- b*v*P_func(u, v, lambda)
  denom <- num + ((1-b)*(1-v)*P_func(1-u, 1-v, lambda))
  return(num/denom)
}
```


## MODEL VARIABLES

```{r}
capital_r <- c(1/4, 1/5, 1/6, 1/7, 1/9, 1/15)
capital_r

n <- c(13, 8, 5, 3, 2, 1)

times <- seq(from=0,to=150,by=1)
xstart <- c(0.4, 0.3, 0.2, 0.1, 0.05, 0.01)
```

## Null Model
b=0.5, lambda=0

```{r}
null_model <- function (t, x, parms) {
  x0 <- 0.5
  x1 <- x[1]
  x2 <- x[2]
  x3 <- x[3]
  x4 <- x[4]
  x5 <- x[5]
  x6 <- x[6]
  cap_r <- parms$cap_r
  L <- length(cap_r)
  cap_n <- parms$cap_n
  low_r <- vector(mode="numeric", length=length(cap_r))
  for(j in seq_along(cap_r)){
  products <- vector(mode="numeric", length=length((j+1):L))
  for(k in seq_along(products)){
    products[k] <- cap_r[j+k]*n[j+k]
    }
  num <- sum(products)
  denom <- cap_r[j] * n[j]
  final <- num/denom
  low_r[j] <- final
  }
  0 -> low_r[is.na(low_r)]
  dx1dt <- cap_r[1] * (1 + low_r[1]) * (x0 - x1)
  dx2dt <- cap_r[2] * (1 + low_r[2]) * (x1 - x2)
  dx3dt <- cap_r[3] * (1 + low_r[3]) * (x2 - x3)
  dx4dt <- cap_r[4] * (1 + low_r[4]) * (x3 - x4)
  dx5dt <- cap_r[5] * (1 + low_r[5]) * (x4 - x5)
  dx6dt <- cap_r[6] * (1 + low_r[6]) * (x5 - x6)
  derivs <- c(dx1dt, dx2dt, dx3dt, dx4dt, dx5dt, dx6dt)
  list(derivs)
}

param_null <- list(cap_r = capital_r, cap_n = n)

null_model <- ode(y = xstart, times = times, func = null_model, parms = param_null)

null_model <- as.data.frame(null_model)
names(null_model) <- c("time", "undergrad", "grad", "postdoc", "tenure track", "tenured prof", "full prof")

null_data <- melt(null_model, id.vars = "time")
head(null_data)
null_graph <- null_data %>% ggplot(aes(x=time, y=value, color=variable)) +
  geom_line() +
  labs(x="Time", y="Fraction Women", color="Levels") +
  theme_bw()
null_graph

pdf("null_model.jpg") 
null_graph
dev.off() 
```


## HOMOPHILY FREE MODEL
lambda=0, set b here

```{r}
model_homo_free <- function (t, x, parms) {
  x0 <- 0.5
  x1 <- x[1]
  x2 <- x[2]
  x3 <- x[3]
  x4 <- x[4]
  x5 <- x[5]
  x6 <- x[6]
  cap_r <- parms$cap_r
  L <- length(cap_r)
  cap_n <- parms$cap_n
  b <- parms$bias
  low_r <- vector(mode="numeric", length=length(cap_r))
  for(j in seq_along(cap_r)){
  products <- vector(mode="numeric", length=length((j+1):L))
  for(k in seq_along(products)){
    products[k] <- cap_r[j+k]*n[j+k]
  }
  num <- sum(products)
  denom <- cap_r[j] * n[j]
  final <- num/denom
  low_r[j] <- final
  }
  0 -> low_r[is.na(low_r)]
  dx1dt <- (cap_r[1]*(1+low_r[1])*(b*x0)/((b*x0)+((1-b)*(1-x0)))) - 
    (cap_r[1]*x1) - ((cap_r[1]*low_r[1]*(b*x1))/((b*x1)+((1-b)*(1-x1))))
  dx2dt <- (cap_r[2]*(1+low_r[2])*(b*x1)/((b*x1)+((1-b)*(1-x1)))) - 
    (cap_r[2]*x2) - ((cap_r[2]*low_r[2]*(b*x2))/((b*x2)+((1-b)*(1-x2))))
  dx3dt <- (cap_r[3]*(1+low_r[3])*(b*x2)/((b*x2)+((1-b)*(1-x2)))) - 
    (cap_r[3]*x3) - ((cap_r[3]*low_r[3]*(b*x3))/((b*x3)+((1-b)*(1-x3))))
  dx4dt <- (cap_r[4]*(1+low_r[4])*(b*x3)/((b*x3)+((1-b)*(1-x3)))) - 
    (cap_r[4]*x4) - ((cap_r[4]*low_r[4]*(b*x4))/((b*x4)+((1-b)*(1-x4))))
  dx5dt <- (cap_r[5]*(1+low_r[5])*(b*x4)/((b*x4)+((1-b)*(1-x4)))) - 
    (cap_r[5]*x5) - ((cap_r[5]*low_r[5]*(b*x5))/((b*x5)+((1-b)*(1-x5))))
  dx6dt <- (cap_r[6]*(1+low_r[6])*(b*x5)/((b*x5)+((1-b)*(1-x5)))) - 
    (cap_r[6]*x6) - ((cap_r[6]*low_r[6]*(b*x6))/((b*x6)+((1-b)*(1-x6))))
  derivs <- c(dx1dt, dx2dt, dx3dt, dx4dt, dx5dt, dx6dt)
  list(derivs)
}

b <- 0.35

param_homo <- list(cap_r = capital_r, cap_n = n, bias = b)

homo_free <- ode(y = xstart, times = times, func = model_homo_free, parms = param_homo)

homo_free <- as.data.frame(homo_free)
names(homo_free) <- c("time", "undergrad", "grad", "postdoc", "tenure track", "tenured prof", "full prof")

homo_free_data <- melt(homo_free, id.vars = "time")

homo_free_data %>% ggplot(aes(x=time, y=value, color=variable)) +
  geom_line() +
  labs(x="Time", y="Fraction Women", color="Levels") +
  theme_bw()
```

## HOMOPHILY FREE, graph multiple bias levels
```{r}
biases_to_test <- c(0.35, 0.49, 0.51, 0.65)

varied_biases_data <- vector(mode="list", length=length(biases_to_test))
names(varied_biases_data) <- paste0("b=", biases_to_test)
for(i in seq_along(biases_to_test)){
  bias_i <- biases_to_test[i]
  parameter <- list(cap_r = capital_r, cap_n = n, bias = bias_i)
  ode_solve <- ode(y = xstart, times = times, func = model_homo_free, parms = parameter)
  ode_solve <- as.data.frame(ode_solve)
  names(ode_solve) <- c("time", "undergrad", "grad", "postdoc", "tenure track", "tenured prof", "full prof")
  data <- melt(ode_solve, id.vars = "time")
  data <- data %>% mutate(b = bias_i)
  varied_biases_data[[i]] <- data
}

data_collected <- do.call(rbind, varied_biases_data)

data_collected %>% ggplot(aes(x=time, y=value, color=variable)) + 
  geom_line() +
  labs(x="Time", y="Fraction Women", color="Levels", title = "Homophily-Free Model") +
  theme_bw() +
  facet_wrap(~ b, labeller = label_both)
```


## BIAS FREE (b=0.5)
set lambda
if I can get this to work right, we can vary bias and lambda over time/level (ambitious women at the top are less susceptible to homophily, bias may increase at higher levels as women approach motherhood, etc) since the parameter input is a list, and I can add the values
```{r}
model_bias_free <- function (t, x, parms) {
  x0 <- 0.5
  x1 <- x[1]
  x2 <- x[2]
  x3 <- x[3]
  x4 <- x[4]
  x5 <- x[5]
  x6 <- x[6]
  cap_r <- parms$cap_r
  L <- length(cap_r)
  cap_n <- parms$cap_n
  lambda <- parms$lambda
  low_r <- vector(mode="numeric", length=length(cap_r))
  for(j in seq_along(cap_r)){
  products <- vector(mode="numeric", length=length((j+1):L))
  for(k in seq_along(products)){
    products[k] <- cap_r[j+k]*cap_n[j+k]
  }
  num <- sum(products)
  denom <- cap_r[j] * cap_n[j]
  final <- num/denom
  low_r[j] <- final
  }
  0 -> low_r[is.na(low_r)]
  dx1dt <- cap_r[1]*(((1+low_r[1])*F0_func(x1, x0, lambda)) - x1 - (low_r[1]*F0_func(x2, x1, lambda)))
  dx2dt <- cap_r[2]*(((1+low_r[2])*F0_func(x2, x1, lambda)) - x2 - (low_r[2]*F0_func(x3, x2, lambda)))
  dx3dt <- cap_r[3]*(((1+low_r[3])*F0_func(x3, x2, lambda)) - x3 - (low_r[3]*F0_func(x4, x3, lambda)))
  dx4dt <- cap_r[4]*(((1+low_r[4])*F0_func(x4, x3, lambda)) - x4 - (low_r[4]*F0_func(x5, x4, lambda)))
  dx5dt <- cap_r[5]*(((1+low_r[5])*F0_func(x5, x4, lambda)) - x5 - (low_r[5]*F0_func(x6, x5, lambda)))
  dx6dt <- cap_r[6]*F0_func(x6, x5, lambda) - cap_r[6]*x6
  derivs <- c(dx1dt, dx2dt, dx3dt, dx4dt, dx5dt, dx6dt)
  list(derivs)
}

lambda <- 5

param_bias <- list(cap_r = capital_r, cap_n = n, lambda = lambda)

bias_free <- ode(y = xstart, times = times, func = model_bias_free, parms = param_bias)

bias_free <- as.data.frame(bias_free)
names(bias_free) <- c("time", "undergrad", "grad", "postdoc", "tenure track", "tenured prof", "full prof")
head(bias_free)

bias_free_data <- melt(bias_free, id.vars = "time")
head(bias_free_data)
bias_free_data %>% ggplot(aes(x=time, y=value, color=variable)) +
  geom_line() +
  labs(x="Time", y="Fraction Women", color="Levels") +
  theme_bw()
```

## BIAS FREE, graph multiple homophily levels
```{r}
homophily_to_test <- c(2, 3, 4.5, 5)
homophily_to_test
varied_homo_data <- vector(mode="list", length=length(homophily_to_test))
names(varied_homo_data) <- paste0("lambda=", homophily_to_test)
for(i in seq_along(homophily_to_test)){
  lambda <- homophily_to_test[i]
  param_bias <- list(cap_r = capital_r, cap_n = n, lambda = lambda)
  sol <- ode(y = xstart, times = times, func = model_bias_free, parms = param_bias)
  sol <- as.data.frame(sol)
  names(sol) <- c("time", "undergrad", "grad", "postdoc", "tenure track", "tenured prof", "full prof")
  dat <- melt(sol, id.vars = "time")
  dat <- dat %>% mutate(homophily=lambda)
  varied_homo_data[[i]] <- dat
}

bias_free_collected <- do.call(rbind, varied_homo_data)

bias_free_collected %>% ggplot(aes(x=time, y=value, color=variable)) + 
  geom_line() +
  labs(x="Time", y="Fraction Women", color="Levels", title = "Bias-Free Model") +
  theme_bw() +
  facet_wrap(~ homophily, labeller = label_both)
```

# FULL MODEL

```{r}
full_model <- function (t, x, parms) {
  x0 <- 0.5
  x1 <- x[1]
  x2 <- x[2]
  x3 <- x[3]
  x4 <- x[4]
  x5 <- x[5]
  x6 <- x[6]
  cap_r <- parms$cap_r
  L <- length(cap_r)
  cap_n <- parms$cap_n
  lambda <- parms$lambda
  b <- parms$bias
  low_r <- vector(mode="numeric", length=length(cap_r))
  for(j in seq_along(cap_r)){
  products <- vector(mode="numeric", length=length((j+1):L))
  for(k in seq_along(products)){
    products[k] <- cap_r[j+k]*cap_n[j+k]
    }
  num <- sum(products)
  denom <- cap_r[j] * cap_n[j]
  final <- num/denom
  low_r[j] <- final
  }
  0 -> low_r[is.na(low_r)]
  dx1dt <- (cap_r[1]*(1+low_r[1])*F_func(u = x1, v = x0, lambda = lambda, b = b)) - (cap_r[1]*x1) - 
    (cap_r[1]*low_r[1]*F_func(u = x2, v = x1, lambda = lambda, b = b))
  dx2dt <- (cap_r[2]*(1+low_r[2])*F_func(u = x2, v = x1, lambda = lambda, b = b)) - (cap_r[2]*x2) - 
    (cap_r[2]*low_r[2]*F_func(u = x3, v = x2, lambda = lambda, b = b))
  dx3dt <- (cap_r[3]*(1+low_r[3])*F_func(u = x3, v = x2, lambda = lambda, b = b)) - (cap_r[3]*x3) - 
    (cap_r[3]*low_r[3]*F_func(u = x4, v = x3, lambda = lambda, b = b))
  dx4dt <- (cap_r[4]*(1+low_r[4])*F_func(u = x4, v = x3, lambda = lambda, b = b)) - (cap_r[4]*x4) - 
    (cap_r[4]*low_r[4]*F_func(u = x5, v = x4, lambda = lambda, b = b))
  dx5dt <- (cap_r[5]*(1+low_r[5])*F_func(u = x5, v = x4, lambda = lambda, b = b)) - (cap_r[5]*x5) - 
    (cap_r[5]*low_r[5]*F_func(u = x6, v = x5, lambda = lambda, b = b))
  dx6dt <- (cap_r[6]*F_func(u = x6, v = x5, lambda = lambda, b = b)) - (cap_r[6]*x6)
  derivs <- c(dx1dt, dx2dt, dx3dt, dx4dt, dx5dt, dx6dt)
  list(derivs)
}

lambda <- 3

b <- 0.3

params <- list(cap_r = capital_r, cap_n = n, lambda = lambda, bias = b)

full_output <- ode(y = xstart, times = times, func = full_model, parms = params)

diagnostics(full_output)

full_output <- as.data.frame(full_output)
names(full_output) <- c("time", "undergrad", "grad", "postdoc", "tenure track", "tenured prof", "full prof")

full_data <- melt(full_output, id.vars = "time")
head(full_data)
full_data %>% ggplot(aes(x=time, y=value, color=variable)) +
  geom_line() +
  labs(x="Time", y="Fraction Women", color="Levels") +
  theme_bw()
```


## FULL MODEL: graph multiple parameter values
```{r}
bias_test <- c(0.3, 0.3, 0.7, 0.7)
homophily_test <- c(2, 4, 2, 4)
trials <- if(length(bias_test) == length(homophily_test)){
  length(bias_test)
  } else{
  0
}

varied_full_data <- vector(mode="list", length=trials)
names(varied_full_data)
length(varied_full_data)
for(i in seq_len(trials)){
  bias_name <- paste0("b=", bias_test[i])
  homo_name <- paste0("l=", homophily_test[i])
  names(varied_full_data)[i] <- paste(bias_name, homo_name, sep = ",")
}
names(varied_full_data)
for(i in seq_len(trials)){
  b <- bias_test[i]
  l <- homophily_test[i]
  params <- list(cap_r = capital_r, cap_n = n, lambda = lambda, bias = b)
  dat <- ode(y = xstart, times = times, func = full_model, parms = params)
  dat <- as.data.frame(dat)
  names(dat) <- c("time", "undergrad", "grad", "postdoc", "tenure track", "tenured prof", "full prof")
  dat_use <- melt(dat, id.vars = "time")
  dat_use <- dat_use %>% mutate(bias=b, homophily=l)
  varied_full_data[[i]] <- dat_use
}

entire_full_model <- do.call(rbind, varied_full_data)
head(entire_full_model)

entire_full_model %>% ggplot(aes(x=time, y=value, color=variable)) + 
  geom_line() +
  theme_bw() +
  facet_grid(bias ~ homophily, labeller = label_both) +
  labs(x="Time", y="Fraction Women", color="Levels", title = "Full Model")
```

