---
title: "gender bias data"
author: "Anna Flieder"
date: "11/26/2021"
output: html_document
---

```{r, warning=FALSE, message=FALSE}
library(tidyverse)
library(lubridate)
library(knitr)
library(reshape2)
library(car)
library(deSolve)
library(readxl)
library(minpack.lm)
library(FME)
library(gridExtra)
```


2012-2020
```{r}
years1 <- 2012:2020

data1 <- vector(mode="list", length=length(years1))
names(data1) <- years1

for(i in seq_along(data1)){
  name <- paste0("data/s", years1[i], "_is.csv")
  read_data <- read.csv(name)
  read_data <- read_data %>% mutate(year=years1[i])
  data1[[i]] <- as.data.frame(read_data)
}


clean_data1 <- function(x){
 portion <- x %>% select(year, UNITID, ARANK, HRTOTLT, HRTOTLM, HRTOTLW) %>% filter(ARANK %in% 1:5)
 summary <- portion %>% group_by(ARANK) %>% summarize(fraction=sum(HRTOTLW)/sum(HRTOTLT))
 fin <- as.data.frame(summary)
 fin <- fin %>% mutate(year = portion$year[1])
 return(as.data.frame(fin))
}

final_data1 <- lapply(data1, clean_data1)

final1 <- do.call(rbind, final_data1)

rankings <- c("full professor", "associate professor", "assistant professor", "instructor", "lecturer")
for(i in seq_along(unique(final1$ARANK))){
  condition <- final1$ARANK == i
  final1[condition, "ARANK"] <- rankings[i]
}

final1_wide <- dcast(final1, year ~ ARANK, value.var="fraction")
col_order <- c("year", "lecturer", "instructor", "assistant professor", "associate professor", "full professor")
final1_wide <- final1_wide[ , col_order]
names(final1_wide)[1] <- "time"
final1_wide

final1_vert <- final1_wide %>% melt(id.vars="time")
final1_vert
```

1999, 2001-2011
```{r}
years2 <- c(1999, 2001:2011)
data2 <- vector(mode="list", length=length(years2))
names(data2) <- years2

for(i in seq_along(data2)){
  name <- paste0("data/s", years2[i], "_f.csv")
  read_data <- read.csv(name)
  read_data <- read_data %>% mutate(year=years2[i])
  data2[[i]] <- as.data.frame(read_data)
}


data2.1 <- data2[9:12]
data2.2 <- data2[5:8]
data2.3 <- data2[1:4]

#hierarchy levels in the data
full_prof <- c(1, 8, 15)
assoc_prof <- c(2, 9, 16)
assist_prof <- c(3, 10, 17)
inst <- c(4, 11, 18)
lect <- c(5, 12, 19)
ranks_to_pull <- c(full_prof, assoc_prof, assist_prof, inst, lect)

clean_data2.1 <- function(x){
  names(x) <- toupper(names(x))
  dat <- x %>% select(YEAR, UNITID, ARANK, HRTOTLT, HRTOTLM, HRTOTLW) %>% filter(ARANK %in% ranks_to_pull)
  is_full_prof <- dat$ARANK %in% full_prof
  is_assoc_prof <- dat$ARANK %in% assoc_prof
  is_assist_prof <- dat$ARANK %in% assist_prof
  is_inst <- dat$ARANK %in% inst
  is_lect <- dat$ARANK %in% lect
  dat[is_full_prof, "ARANK"] <- "full professor"
  dat[is_assoc_prof, "ARANK"] <- "associate professor"
  dat[is_assist_prof, "ARANK"] <- "assistant professor"
  dat[is_inst, "ARANK"] <- "instructor"
  dat[is_lect, "ARANK"] <- "lecturer"
  sums <- dat %>% group_by(ARANK) %>% summarize(fraction=sum(HRTOTLW)/sum(HRTOTLT)) %>% mutate(year=dat$YEAR[1])
  sums <- as.data.frame(sums)
  return(sums)
}

#STAFF24=GRAND TOTAL
#STAFF15=TOTAL MEN
#STAFF16=TOTAL WOMEN
clean_data2.2 <- function(x){
  names(x) <- toupper(names(x))
  dat <- x %>% select(YEAR, UNITID, ARANK, STAFF24, STAFF15, STAFF16) %>% filter(ARANK %in% ranks_to_pull)
  is_full_prof <- dat$ARANK %in% full_prof
  is_assoc_prof <- dat$ARANK %in% assoc_prof
  is_assist_prof <- dat$ARANK %in% assist_prof
  is_inst <- dat$ARANK %in% inst
  is_lect <- dat$ARANK %in% lect
  dat[is_full_prof, "ARANK"] <- "full professor"
  dat[is_assoc_prof, "ARANK"] <- "associate professor"
  dat[is_assist_prof, "ARANK"] <- "assistant professor"
  dat[is_inst, "ARANK"] <- "instructor"
  dat[is_lect, "ARANK"] <- "lecturer"
  sums <- dat %>% group_by(ARANK) %>% summarize(fraction=sum(STAFF16)/sum(STAFF24)) %>% mutate(year=dat$YEAR[1])
  sums <- as.data.frame(sums)
  return(sums)
}

#STAFF16=TOTAL WOMEN
#STAFF15=TOTAL MEN
clean_data2.3 <- function(x){
  names(x) <- toupper(names(x))
  dat <- x %>% select(YEAR, UNITID, ARANK, STAFF15, STAFF16) %>% filter(ARANK %in% ranks_to_pull)
  is_full_prof <- dat$ARANK %in% full_prof
  is_assoc_prof <- dat$ARANK %in% assoc_prof
  is_assist_prof <- dat$ARANK %in% assist_prof
  is_inst <- dat$ARANK %in% inst
  is_lect <- dat$ARANK %in% lect
  dat[is_full_prof, "ARANK"] <- "full professor"
  dat[is_assoc_prof, "ARANK"] <- "associate professor"
  dat[is_assist_prof, "ARANK"] <- "assistant professor"
  dat[is_inst, "ARANK"] <- "instructor"
  dat[is_lect, "ARANK"] <- "lecturer"
  dat <- dat %>% mutate(TOTAL = STAFF15 + STAFF16)
  dat <- as.data.frame(dat)
  sums <- dat %>% group_by(ARANK) %>% summarize(fraction=sum(STAFF16)/sum(TOTAL)) %>% mutate(year=dat$YEAR[1])
  sums <- as.data.frame(sums)
  return(sums)
}


final_data2.1 <- lapply(data2.1, clean_data2.1)

final_data2.2 <- lapply(data2.2, clean_data2.2)

final_data2.3 <- lapply(data2.3, clean_data2.3)

final_data2 <- c(final_data2.1, final_data2.2, final_data2.3)
final2 <- do.call(rbind, final_data2)
final2_wide <- dcast(final2, year ~ ARANK, value.var="fraction")
final2_wide <- final2_wide[ ,col_order]
names(final2_wide)[1] <- "time"
final2_wide

final2_vert <- final2_wide %>% melt(id.vars="time")
final2_vert
```

1997, 1995
-STAFF15=total men
-STAFF16=total women
-need to manually compute total
-part F = full-time faculty by academic rank and tenure
1993
-staff15=total men
-staff16=total women
-need to manually compute total
```{r}
years3.1 <- c(1995, 1997)

data3.1 <- vector(mode="list", length=length(years3.1))
names(data3.1) <- years3.1
for(i in seq_along(data3.1)){
  name <- paste0("data/s", years3.1[i], "_s.csv")
  read_data <- read.csv(name)
  read_data <- read_data %>% mutate(year=years3.1[i])
  data3.1[[i]] <- as.data.frame(read_data)
}

full_profs3 <- c(86, 93, 100)
assoc_profs3 <- c(87, 94, 101)
assist_profs3 <- c(88, 95, 102)
inst3 <- c(89, 96, 103)
lect3 <- c(90, 97, 104)
ranks_to_pull3 <- c(full_profs3, assoc_profs3, assist_profs3, inst3, lect3)

clean_data3.1 <- function(x){
  names(x) <- toupper(names(x))
  dat <- x %>% select(YEAR, UNITID, PART, LINE, STAFF15, STAFF16) %>% filter(PART == "F", LINE %in% ranks_to_pull3)
  is_full_prof <- dat$LINE %in% full_profs3
  is_assoc_prof <- dat$LINE %in% assoc_profs3
  is_assist_prof <- dat$LINE %in% assist_profs3
  is_inst <- dat$LINE %in% inst3
  is_lect <- dat$LINE %in% lect3
  dat[is_full_prof, "LINE"] <- "full professor"
  dat[is_assoc_prof, "LINE"] <- "associate professor"
  dat[is_assist_prof, "LINE"] <- "assistant professor"
  dat[is_inst, "LINE"] <- "instructor"
  dat[is_lect, "LINE"] <- "lecturer"
  dat <- dat %>% mutate(TOTAL = STAFF15 + STAFF16)
  dat <- as.data.frame(dat)
  sums <- dat %>% group_by(LINE) %>% summarize(fraction=sum(STAFF16)/sum(TOTAL)) %>% mutate(year=dat$YEAR[1])
  sums <- as.data.frame(sums)
  return(sums)
}

final_data3.1 <- lapply(data3.1, clean_data3.1)

final3.1 <- do.call(rbind, final_data3.1)

final3.1
names(final3.1) <- c("ARANK", "fraction", "year")

#1993
years3.2 <- 1993

full_profs3.2 <- c(78, 85, 92)
assoc_profs3.2 <- c(79, 86, 93)
assist_profs3.2 <- c(80, 87, 94)
inst3.2 <- c(81, 88, 95)
lect3.2 <- c(82, 89, 96)
ranks_to_pull3.2 <- c(full_profs3.2, assoc_profs3.2, assist_profs3.2, inst3.2, lect3.2)

data3.2 <- read.csv("data/s1993_abcef.csv")

final_data3.2 <- data3.2 %>% select(unitid, line, staff15, staff16) %>% filter(line %in% ranks_to_pull3.2) %>%
  mutate(total= staff15+staff16)
is_full_prof <- final_data3.2$line %in% full_profs3.2
is_assoc_prof <- final_data3.2$line %in% assoc_profs3.2
is_assist_prof <- final_data3.2$line %in% assist_profs3.2
is_inst <- final_data3.2$line %in% inst3.2
is_lect <- final_data3.2$line %in% lect3.2
final_data3.2[is_full_prof, "line"] <- "full professor"
final_data3.2[is_assoc_prof, "line"] <- "associate professor"
final_data3.2[is_assist_prof, "line"] <- "assistant professor"
final_data3.2[is_inst, "line"] <- "instructor"
final_data3.2[is_lect, "line"] <- "lecturer"
final_data3.2

final3.2 <- final_data3.2 %>% group_by(line) %>% summarize(fraction= sum(staff16)/sum(total)) %>% mutate(year=years3.2)
final3.2 <- as.data.frame(final3.2)
names(final3.2) <- c("ARANK", "fraction", "year")
final3.2

final3 <- rbind(final3.1, final3.2)
final3

final3_wide <- dcast(final3, year ~ ARANK, value.var="fraction")
final3_wide <- final3_wide[ , col_order]
names(final3_wide)[1] <- "time"
final3_wide

final3_vert <- final3_wide %>% melt(id.vars="time")
final3_vert <- final3_vert %>% arrange(time)
final3_vert
```


ENTIRE DATA
```{r}
wide_data <- rbind(final1_wide, final2_wide, final3_wide)
wide_data <- wide_data %>% arrange(time)
wide_data <- as.data.frame(wide_data)
wide_data
write.csv(wide_data, "clean_wide_data.csv")

vert_data <- rbind(final1_vert, final2_vert, final3_vert)
vert_data <- vert_data %>% arrange(time)
vert_data
write.csv(vert_data, "clean_vert_data.csv")

years <- unique(wide_data$time)
years

vert_data %>% ggplot(aes(x=time, y=value, color=variable)) +
  geom_line() +
  labs(x="Year", y="Fraction Women", color="Levels", title="All Universities") +
  theme_bw()
```


## COLUMBIA DATA
columbia UNITID = 190150

2012-2020
```{r}
clean_data1cu <- function(x){
 portion <- x %>% select(year, UNITID, ARANK, HRTOTLT, HRTOTLM, HRTOTLW) %>% filter(ARANK %in% 1:5, UNITID == 190150)
 summary <- portion %>% group_by(ARANK) %>% summarize(fraction=sum(HRTOTLW)/sum(HRTOTLT))
 fin <- as.data.frame(summary)
 fin <- fin %>% mutate(year = portion$year[1])
 return(as.data.frame(fin))
}

final_data1cu <- lapply(data1, clean_data1cu)

final1cu <- do.call(rbind, final_data1cu)

for(i in seq_along(unique(final1$ARANK))){
  condition <- final1cu$ARANK == i
  final1cu[condition, "ARANK"] <- rankings[i]
}

final1cu_wide <- dcast(final1cu, year ~ ARANK, value.var="fraction")
final1cu_wide <- final1cu_wide[ , col_order]
names(final1cu_wide)[1] <- "time"
final1cu_wide

final1cu_vert <- final1cu_wide %>% melt(id.vars="time")
final1cu_vert
```

1999, 2001-2011
```{r}
clean_data2.1cu <- function(x){
  names(x) <- toupper(names(x))
  dat <- x %>% select(YEAR, UNITID, ARANK, HRTOTLT, HRTOTLM, HRTOTLW) %>% filter(ARANK %in% ranks_to_pull, UNITID == 190150)
  is_full_prof <- dat$ARANK %in% full_prof
  is_assoc_prof <- dat$ARANK %in% assoc_prof
  is_assist_prof <- dat$ARANK %in% assist_prof
  is_inst <- dat$ARANK %in% inst
  is_lect <- dat$ARANK %in% lect
  dat[is_full_prof, "ARANK"] <- "full professor"
  dat[is_assoc_prof, "ARANK"] <- "associate professor"
  dat[is_assist_prof, "ARANK"] <- "assistant professor"
  dat[is_inst, "ARANK"] <- "instructor"
  dat[is_lect, "ARANK"] <- "lecturer"
  sums <- dat %>% group_by(ARANK) %>% summarize(fraction=sum(HRTOTLW)/sum(HRTOTLT)) %>% mutate(year=dat$YEAR[1])
  sums <- as.data.frame(sums)
  return(sums)
}

clean_data2.2cu <- function(x){
  names(x) <- toupper(names(x))
  dat <- x %>% select(YEAR, UNITID, ARANK, STAFF24, STAFF15, STAFF16) %>% filter(ARANK %in% ranks_to_pull, UNITID == 190150)
  is_full_prof <- dat$ARANK %in% full_prof
  is_assoc_prof <- dat$ARANK %in% assoc_prof
  is_assist_prof <- dat$ARANK %in% assist_prof
  is_inst <- dat$ARANK %in% inst
  is_lect <- dat$ARANK %in% lect
  dat[is_full_prof, "ARANK"] <- "full professor"
  dat[is_assoc_prof, "ARANK"] <- "associate professor"
  dat[is_assist_prof, "ARANK"] <- "assistant professor"
  dat[is_inst, "ARANK"] <- "instructor"
  dat[is_lect, "ARANK"] <- "lecturer"
  sums <- dat %>% group_by(ARANK) %>% summarize(fraction=sum(STAFF16)/sum(STAFF24)) %>% mutate(year=dat$YEAR[1])
  sums <- as.data.frame(sums)
  return(sums)
}

clean_data2.3cu <- function(x){
  names(x) <- toupper(names(x))
  dat <- x %>% select(YEAR, UNITID, ARANK, STAFF15, STAFF16) %>% filter(ARANK %in% ranks_to_pull, UNITID == 190150)
  is_full_prof <- dat$ARANK %in% full_prof
  is_assoc_prof <- dat$ARANK %in% assoc_prof
  is_assist_prof <- dat$ARANK %in% assist_prof
  is_inst <- dat$ARANK %in% inst
  is_lect <- dat$ARANK %in% lect
  dat[is_full_prof, "ARANK"] <- "full professor"
  dat[is_assoc_prof, "ARANK"] <- "associate professor"
  dat[is_assist_prof, "ARANK"] <- "assistant professor"
  dat[is_inst, "ARANK"] <- "instructor"
  dat[is_lect, "ARANK"] <- "lecturer"
  dat <- dat %>% mutate(TOTAL = STAFF15 + STAFF16)
  dat <- as.data.frame(dat)
  sums <- dat %>% group_by(ARANK) %>% summarize(fraction=sum(STAFF16)/sum(TOTAL)) %>% mutate(year=dat$YEAR[1])
  sums <- as.data.frame(sums)
  return(sums)
}


final_data2.1cu <- lapply(data2.1, clean_data2.1cu)

final_data2.2cu <- lapply(data2.2, clean_data2.2cu)

final_data2.3cu <- lapply(data2.3, clean_data2.3cu)

final_data2cu <- c(final_data2.1cu, final_data2.2cu, final_data2.3cu)
final2cu <- do.call(rbind, final_data2cu)
final2cu_wide <- dcast(final2cu, year ~ ARANK, value.var="fraction")
final2cu_wide <- final2cu_wide[ ,col_order]
names(final2cu_wide)[1] <- "time"
final2cu_wide

final2cu_vert <- final2cu_wide %>% melt(id.vars="time")
final2cu_vert
```

1997, 1995
```{r}
clean_data3.1cu <- function(x){
  names(x) <- toupper(names(x))
  dat <- x %>% select(YEAR, UNITID, PART, LINE, STAFF15, STAFF16) %>% 
    filter(PART == "F", LINE %in% ranks_to_pull3, UNITID == 190150)
  is_full_prof <- dat$LINE %in% full_profs3
  is_assoc_prof <- dat$LINE %in% assoc_profs3
  is_assist_prof <- dat$LINE %in% assist_profs3
  is_inst <- dat$LINE %in% inst3
  is_lect <- dat$LINE %in% lect3
  dat[is_full_prof, "LINE"] <- "full professor"
  dat[is_assoc_prof, "LINE"] <- "associate professor"
  dat[is_assist_prof, "LINE"] <- "assistant professor"
  dat[is_inst, "LINE"] <- "instructor"
  dat[is_lect, "LINE"] <- "lecturer"
  dat <- dat %>% mutate(TOTAL = STAFF15 + STAFF16)
  dat <- as.data.frame(dat)
  sums <- dat %>% group_by(LINE) %>% summarize(fraction=sum(STAFF16)/sum(TOTAL)) %>% mutate(year=dat$YEAR[1])
  sums <- as.data.frame(sums)
  return(sums)
}

final_data3.1cu <- lapply(data3.1, clean_data3.1cu)

final3.1cu <- do.call(rbind, final_data3.1cu)
names(final3.1cu) <- c("ARANK", "fraction", "year")
final3.1cu

#1993
final_data3.2cu <- data3.2 %>% select(unitid, line, staff15, staff16) %>% 
  filter(line %in% ranks_to_pull3.2, unitid == 190150) %>% mutate(total= staff15+staff16)
is_full_prof <- final_data3.2cu$line %in% full_profs3.2
is_assoc_prof <- final_data3.2cu$line %in% assoc_profs3.2
is_assist_prof <- final_data3.2cu$line %in% assist_profs3.2
is_inst <- final_data3.2cu$line %in% inst3.2
is_lect <- final_data3.2cu$line %in% lect3.2
final_data3.2cu[is_full_prof, "line"] <- "full professor"
final_data3.2cu[is_assoc_prof, "line"] <- "associate professor"
final_data3.2cu[is_assist_prof, "line"] <- "assistant professor"
final_data3.2cu[is_inst, "line"] <- "instructor"
final_data3.2cu[is_lect, "line"] <- "lecturer"
final_data3.2cu

final3.2cu <- final_data3.2cu %>% group_by(line) %>% summarize(fraction= sum(staff16)/sum(total)) %>% mutate(year=years3.2)
final3.2cu <- as.data.frame(final3.2cu)
names(final3.2cu) <- c("ARANK", "fraction", "year")
final3.2cu

final3cu <- rbind(final3.1cu, final3.2cu)
final3cu

final3cu_wide <- dcast(final3cu, year ~ ARANK, value.var="fraction")
final3cu_wide <- final3cu_wide[ , col_order]
names(final3cu_wide)[1] <- "time"
final3cu_wide

final3cu_vert <- final3cu_wide %>% melt(id.vars="time") %>% arrange(time)
final3cu_vert
```


ENTIRE DATA
```{r}
wide_data_cu <- rbind(final1cu_wide, final2cu_wide, final3cu_wide) %>% arrange(time)
wide_data_cu <- as.data.frame(wide_data_cu)
wide_data_cu
write.csv(wide_data_cu, "columbia_clean_wide_data.csv")

vert_data_cu <- rbind(final1cu_vert, final2cu_vert, final3cu_vert) %>% arrange(time)
vert_data_cu <- as.data.frame(vert_data_cu)
vert_data_cu
write.csv(vert_data_cu, "columbia_clean_vert_data.csv")

years_cu <- unique(wide_data_cu$time)
years_cu

vert_data_cu %>% ggplot(aes(x=time, y=value, color=variable)) +
  geom_line() +
  labs(x="Year", y="Fraction Women", color="Levels", title = "Columbia") +
  theme_bw()
```


## BARNARD
barnard UNITID= 189097

2012-2020
```{r}
clean_data1bc <- function(x){
 portion <- x %>% select(year, UNITID, ARANK, HRTOTLT, HRTOTLM, HRTOTLW) %>% filter(ARANK %in% 1:5, UNITID == 189097)
 summary <- portion %>% group_by(ARANK) %>% summarize(fraction=sum(HRTOTLW)/sum(HRTOTLT))
 fin <- as.data.frame(summary)
 fin <- fin %>% mutate(year = portion$year[1])
 return(as.data.frame(fin))
}

final_data1bc <- lapply(data1, clean_data1bc)

final1bc <- do.call(rbind, final_data1bc)

for(i in seq_along(unique(final1$ARANK))){
  condition <- final1bc$ARANK == i
  final1bc[condition, "ARANK"] <- rankings[i]
}

final1bc_wide <- dcast(final1bc, year ~ ARANK, value.var="fraction")
final1bc_wide <- final1bc_wide[ , col_order]
names(final1bc_wide)[1] <- "time"
final1bc_wide

final1bc_vert <- final1bc_wide %>% melt(id.vars="time")
final1bc_vert
```

1999, 2001-2011
```{r}
clean_data2.1bc <- function(x){
  names(x) <- toupper(names(x))
  dat <- x %>% select(YEAR, UNITID, ARANK, HRTOTLT, HRTOTLM, HRTOTLW) %>% filter(ARANK %in% ranks_to_pull, UNITID == 189097)
  is_full_prof <- dat$ARANK %in% full_prof
  is_assoc_prof <- dat$ARANK %in% assoc_prof
  is_assist_prof <- dat$ARANK %in% assist_prof
  is_inst <- dat$ARANK %in% inst
  is_lect <- dat$ARANK %in% lect
  dat[is_full_prof, "ARANK"] <- "full professor"
  dat[is_assoc_prof, "ARANK"] <- "associate professor"
  dat[is_assist_prof, "ARANK"] <- "assistant professor"
  dat[is_inst, "ARANK"] <- "instructor"
  dat[is_lect, "ARANK"] <- "lecturer"
  sums <- dat %>% group_by(ARANK) %>% summarize(fraction=sum(HRTOTLW)/sum(HRTOTLT)) %>% mutate(year=dat$YEAR[1])
  sums <- as.data.frame(sums)
  return(sums)
}

clean_data2.2bc <- function(x){
  names(x) <- toupper(names(x))
  dat <- x %>% select(YEAR, UNITID, ARANK, STAFF24, STAFF15, STAFF16) %>% filter(ARANK %in% ranks_to_pull, UNITID == 189097)
  is_full_prof <- dat$ARANK %in% full_prof
  is_assoc_prof <- dat$ARANK %in% assoc_prof
  is_assist_prof <- dat$ARANK %in% assist_prof
  is_inst <- dat$ARANK %in% inst
  is_lect <- dat$ARANK %in% lect
  dat[is_full_prof, "ARANK"] <- "full professor"
  dat[is_assoc_prof, "ARANK"] <- "associate professor"
  dat[is_assist_prof, "ARANK"] <- "assistant professor"
  dat[is_inst, "ARANK"] <- "instructor"
  dat[is_lect, "ARANK"] <- "lecturer"
  sums <- dat %>% group_by(ARANK) %>% summarize(fraction=sum(STAFF16)/sum(STAFF24)) %>% mutate(year=dat$YEAR[1])
  sums <- as.data.frame(sums)
  return(sums)
}

clean_data2.3bc <- function(x){
  names(x) <- toupper(names(x))
  dat <- x %>% select(YEAR, UNITID, ARANK, STAFF15, STAFF16) %>% filter(ARANK %in% ranks_to_pull, UNITID == 189097)
  is_full_prof <- dat$ARANK %in% full_prof
  is_assoc_prof <- dat$ARANK %in% assoc_prof
  is_assist_prof <- dat$ARANK %in% assist_prof
  is_inst <- dat$ARANK %in% inst
  is_lect <- dat$ARANK %in% lect
  dat[is_full_prof, "ARANK"] <- "full professor"
  dat[is_assoc_prof, "ARANK"] <- "associate professor"
  dat[is_assist_prof, "ARANK"] <- "assistant professor"
  dat[is_inst, "ARANK"] <- "instructor"
  dat[is_lect, "ARANK"] <- "lecturer"
  dat <- dat %>% mutate(TOTAL = STAFF15 + STAFF16)
  dat <- as.data.frame(dat)
  sums <- dat %>% group_by(ARANK) %>% summarize(fraction=sum(STAFF16)/sum(TOTAL)) %>% mutate(year=dat$YEAR[1])
  sums <- as.data.frame(sums)
  return(sums)
}


final_data2.1bc <- lapply(data2.1, clean_data2.1bc)

final_data2.2bc <- lapply(data2.2, clean_data2.2bc)

final_data2.3bc <- lapply(data2.3, clean_data2.3bc)

final_data2bc <- c(final_data2.1bc, final_data2.2bc, final_data2.3bc)
final2bc <- do.call(rbind, final_data2bc)
final2bc_wide <- dcast(final2bc, year ~ ARANK, value.var="fraction")
final2bc_wide <- final2bc_wide[ ,col_order]
names(final2bc_wide)[1] <- "time"
final2bc_wide

final2bc_vert <- final2bc_wide %>% melt(id.vars="time")
final2bc_vert
```

1997, 1995
```{r}
clean_data3.1bc <- function(x){
  names(x) <- toupper(names(x))
  dat <- x %>% select(YEAR, UNITID, PART, LINE, STAFF15, STAFF16) %>% 
    filter(PART == "F", LINE %in% ranks_to_pull3, UNITID == 189097)
  is_full_prof <- dat$LINE %in% full_profs3
  is_assoc_prof <- dat$LINE %in% assoc_profs3
  is_assist_prof <- dat$LINE %in% assist_profs3
  is_inst <- dat$LINE %in% inst3
  is_lect <- dat$LINE %in% lect3
  dat[is_full_prof, "LINE"] <- "full professor"
  dat[is_assoc_prof, "LINE"] <- "associate professor"
  dat[is_assist_prof, "LINE"] <- "assistant professor"
  dat[is_inst, "LINE"] <- "instructor"
  dat[is_lect, "LINE"] <- "lecturer"
  dat <- dat %>% mutate(TOTAL = STAFF15 + STAFF16)
  dat <- as.data.frame(dat)
  sums <- dat %>% group_by(LINE) %>% summarize(fraction=sum(STAFF16)/sum(TOTAL)) %>% mutate(year=dat$YEAR[1])
  sums <- as.data.frame(sums)
  return(sums)
}

final_data3.1bc <- lapply(data3.1, clean_data3.1bc)

final3.1bc <- do.call(rbind, final_data3.1bc)
names(final3.1bc) <- c("ARANK", "fraction", "year")
final3.1bc

#1993
final_data3.2bc <- data3.2 %>% select(unitid, line, staff15, staff16) %>% 
  filter(line %in% ranks_to_pull3.2, unitid == 189097) %>% mutate(total= staff15+staff16)
is_full_prof <- final_data3.2bc$line %in% full_profs3.2
is_assoc_prof <- final_data3.2bc$line %in% assoc_profs3.2
is_assist_prof <- final_data3.2bc$line %in% assist_profs3.2
is_inst <- final_data3.2bc$line %in% inst3.2
is_lect <- final_data3.2bc$line %in% lect3.2
final_data3.2bc[is_full_prof, "line"] <- "full professor"
final_data3.2bc[is_assoc_prof, "line"] <- "associate professor"
final_data3.2bc[is_assist_prof, "line"] <- "assistant professor"
final_data3.2bc[is_inst, "line"] <- "instructor"
final_data3.2bc[is_lect, "line"] <- "lecturer"
final_data3.2bc

final3.2bc <- final_data3.2bc %>% group_by(line) %>% summarize(fraction= sum(staff16)/sum(total)) %>% mutate(year=years3.2)
final3.2bc <- as.data.frame(final3.2bc)
names(final3.2bc) <- c("ARANK", "fraction", "year")
final3.2bc

final3bc <- rbind(final3.1bc, final3.2bc)
final3bc

final3bc_wide <- dcast(final3bc, year ~ ARANK, value.var="fraction")
final3bc_wide <- final3bc_wide[ , col_order]
names(final3bc_wide)[1] <- "time"
final3bc_wide

final3bc_vert <- final3bc_wide %>% melt(id.vars="time") %>% arrange(time)
final3bc_vert
```

ENTIRE DATA
barnard has spotty "instructor" data -- probably because as a college it does not hire many lecturers, so i take it out
```{r}
wide_data_bc <- rbind(final1bc_wide, final2bc_wide, final3bc_wide) %>% 
  arrange(time) %>% select("time", "lecturer", "assistant professor", "associate professor", "full professor")
wide_data_bc <- as.data.frame(wide_data_bc)
wide_data_bc
write.csv(wide_data_bc, "barnard_clean_wide_data.csv")

vert_data_bc <- rbind(final1bc_vert, final2bc_vert, final3bc_vert) %>% 
  arrange(time) %>% filter(variable %in% c("lecturer", "assistant professor", "associate professor", "full professor"))
vert_data_bc <- as.data.frame(vert_data_bc)
vert_data_bc
write.csv(vert_data_bc, "barnard_clean_vert_data.csv")

years_bc <- unique(wide_data_bc$time)
years_bc

vert_data_bc %>% ggplot(aes(x=time, y=value, color=variable)) +
  geom_line() +
  labs(x="Year", y="Fraction Women", color="Levels", title = "Barnard") +
  theme_bw()
```
